Smart Representations: {R}ationality and Evolution in a Richer Environment

### 2016-01-25 paper/paper.tex
Paolo almost final

\begin{abstract}
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this (possibly false)
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals of what it means to make a rational choice. Our
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

@@

Evolutionary game theory has become an established philosophical tool for probing into
conceptual issues whose complexity requires mathematical modelling at a high level of
abstraction.  Abstraction entails simplifying assumptions. Evolutionary game models frequently
simplify reality in two ways that are relevant for present purposes. Firstly, the environment
is represented as a fixed, closed and immutable \emph{stage game}; secondly, the focus of
evolutionary selection is behavior for this stage game alone. In contrast, some argue for
studying instead the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. To show that
there is no tension here, this paper introduces a method for reconciling established notions of
evolutionary game theory with evolutionary selection of choice mechanisms in variable
environments \citep[see related work
by][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}

If agents deal with a rich and variable environment, they face many different choice
situations. A choice mechanism associates decision situations with action choices. A crucial
part of any choice mechanism is a \emph{subjective representation} of the decision situation,
in particular a manner of forming preferences and beliefs about a possibly uncertain world (see
Figure~\ref{fig:ChoiceMechanism} below). This paper asks: which preference and belief
representations are ecologically valuable (i.e., lead to high fitness)? The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. The particular contribution of this paper is
to provide a case, spelled out formally, that questions of preference evolution should take
variability in belief representation into account as well. As a point in case, we demonstrate
that if agents have imprecise probabilistic beliefs \citep[e.g.][]{gardsah82,levi74,walley96},
preference representations in terms of regret can be selected for, even though they may deviate
from the true, objective fitness that evolutionary selection operates on.

@@

Section \ref{sec:rati--subj} reviews different views of rational choice and relates the
approach we take here to the existing literature. Section \ref{sec:basic-notions} introduces
relevant decision- and game-theoretic notions and spells out four choice mechanisms on whose
evolutionary competition we will zoom in Section~\ref{sec:basic-model-1}, which presents the
main results of the paper. Section \ref{sec:extensions} extends the evolutionary analysis by
introducing more choice mechanisms and alternative models of the environment.

\section{Rationality, subjective representation schemes \& meta games}
\label{sec:rati--subj}

The standard textbook definition of \textit{rationality} in economics and decision theory
traces back to the seminal work of \citet{deFinetti37}, \citet{Neumannvon-NeumannMorgenstern1944:Theory-of-Games}
and \citet{Savage1954:The-Foundations}. It says that choice is rational only if it maximizes
(subjective) expected utility.

\begin{definition}[Rational choice from maximization of expected utility]
  \label{def:rationality}
  Let $S$ be a (finite) set of states of the world and $A$ a (finite) set of actions. Given a
  probability distribution $\mu$ over $S$ and a payoff function
  $\pi \mycolon S \times A \rightarrow \mathbb{R} $, an action $a^* \in A$ is a rational choice
  if it maximizes the \emph{(subjective) expected utility}
  $\EU(a) = \sum_{s \in S} \mu(s) \ \pi(s, a)$.
\end{definition}

\noindent Expected utility is subjective in the sense that it is a function of subjective
preferences and subjective beliefs of the decision maker (DM). To wit, a choice can be
rational, i.e., the best choice from the DM's point of view, even if based on peculiar beliefs
and/or aberrant preferences. % (Indeed, while
% \citeauthor{Neumannvon-NeumannMorgenstern1944:Theory-of-Games} assumed that probabilities were
% objectively given, it was \citet{Savage1954:The-Foundations} who famously extended von Neumann
% and Morgenstern's axiomatization to include a subjective notion of probabilistic beliefs on top
% of the subjective notion of preference that von Neumann and Morgenstern started out with.)

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. For example,
\citet{KahnemannTversky1979:Prospect-Theory} famously demonstrated that what appear to be
violations of rationality norms in human decision making can be explained on the assumption
that subjects' beliefs and preferences deviate systematically from the objectively given
parameters in the presented choice task. Similarly for social decision making, including social
preferences such as fairness, reciprocity or similar, allows us to describe as rational
empirically observed behavior, such as in experimental prisoners' dilemmas or public goods
games, that might otherwise appear irrational \citep[e.g.,][]{fehrschmidt99,charrab02}.

The main objection to redescriptionism is that, without additional constraints, the notion of
rationality is likely to collapse, as it seems possible to deem rational almost everything that
is observed, given the freedom to adjust beliefs and preferences at will. \emph{Normativism}
therefore emphasizes that there are many ways in which ascriptions of beliefs and preferences
are constrained by normative considerations of rationality as well. As for beliefs, where
possible, subjective beliefs should reflect objective chance. Where this is impossible,
temporal consistency of choices constrains rational subjective beliefs, else exploitation by
Dutch books is possible. As for preferences, in certain situations it may make sense to
postulate on normative grounds that subjective preferences should be oriented towards tracking
objective fitness. For instance, profit maximization seems a reasonable assumption for
characterizing outcomes in a competitive market because only firms behaving in a manner that is
consistent with profit maximization will survive in the long run \myalert{(Alchian,
  Friedman)}.\todo{@Paolo: please add references}

An alternative view on rationality of choice is \emph{adaptationism}
\citep[e.g.,][]{Anderson1991:Is-human-cognit,ChaterOaksford2000:The-Rational-An,HagenChater2012:Decision-Making}. \todo{more
  references} Adaptationism aims to explain choice behavior by appeal to evolutionary
considerations: DMs have acquired choice mechanisms that have proved to be adpative in a rich
and variable environment. A choice mechanism could be a set of distinct heuristics (the DM's
\emph{adaptive toolbox}) that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to relate to the philosophical debate about the nature of rational choice, we here suggest to
think of a choice mechanism as a map from choice situations to action choices that contains an
explicit level of subjective representation of a decision situation (see
Figure~\ref{fig:ChoiceMechanism}). We are most interested in the question which
\emph{subjective representation schemes}, i.e., general ways of forming preferences and
beliefs, are better than others from an evolutionary point of view.

@@

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = environment, node distance = 2cm] (situation)
    {environment produces random \\ \textbf{decision situation} $d$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = situation] (representation) 
    {\textbf{subjective representation} $s$ of $d$ \\ \textcolor{gray}{[preference, beliefs, \dots]}};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = representation] (action)
    {\textbf{action choice} $a$ based on $s$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = action,node distance = 1.5cm] (payoff) {\textbf{payoff} for $a$ in $d$};

@@

\end{tikzpicture}
  \caption{Schematic representation of action choice. A choice mechanism maps decision
    situation $d$ to action choice $a$ by way of representing $d$ in terms of subjective
    preferences and subjective beliefs.}
  \label{fig:ChoiceMechanism}
\end{figure}

The fitness of a subjective representation scheme is measured in terms of the fitness of a
choice mechanism that contains it, all else equal. The fitness of a choice mechanism $c$ is the
expected payoff of $c$ over arbitrary decision situations $d$:
\begin{align}
  \label{eq:FittnessChoiceMechSolitary}
  F(c) = \int P(d) \  \pi_d(c) \text{ d} d \,,
\end{align}
where $P(d)$ is the probability (density) of $d$ occurring and $\pi_d(c)$ is the payoff of
the act selected by $c$ in $d$. Probability measure $P(d)$ is our model of the statistical
properties of a rich and variable environment.

The same logic also applies to situations of social decision making. A game $g$ is also a
decision situation, albeit a more complex one. A choice mechanism $c$ maps $g$ onto an action
choice in $g$. The fitness of $c$ is then:\footnote{We assume for simplicity of exposition that
  all games with positive occurrence probability are two player games.}
\begin{align}
  \label{eq:FittnessChoiceMechGame}
  F(c) = \int \int P(g) \  P(c') \  \pi_g(c,c') \text{ d} g \,,
\end{align}
where $P(g)$ is the probability (density) of game $g$, $P(c')$ is the probability of
encountering a co-player who uses choice mechanism $c'$ and $\pi_g(c,c')$ is the payoff for
a $c$-player in $g$ against the $c'$-player, given by the payoff in $g$ of the actions that $c$
and $c'$ select in $g$. For the most part, we will focus on the more complex case of social
decision making here. Results about solitary decision making follow from parallel arguments
(see Section~\ref{sec:extensions}).

\emph{Meta games} are abstract models for the evolutionary competition between choice
mechanisms in social decision making contexts. The expected payoff for a $c$-player in a random
encounter with a $c'$-player is the average of payoffs obtained in a random game $g$ weighted
by $g$'s occurrence probability:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P(g) \  \pi_g(c,c') \text{ d} g \,.
\end{align}

\begin{definition}[Meta game]
  \label{def:MetaGame}
  Let $C$ be a set of choice mechanisms and let $P(g)$ give the occurrence probability of game
  $g$ in the environment. The meta game that captures selection of choice mechanisms in this
  environment is the one-population evolutionary game with strategies $C$ and payoff function
  $\pi \mycolon C \times C \rightarrow \mathbb{R}$ such that
  $\pi(c,c') = F(c,c') = \int P(g) \  \pi_g(c,c') \text{ d} g$.
\end{definition}

Standard notions of evolutionary game theory can be applied to meta games as well. A choice
principle $c$ is a strict Nash equilibrium if $F(c,c) > F(c',c)$ for all $c'$; it is
evolutionarily stable if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$
and $F(c,c') > F(c',c')$. It is neutrally stable if for all $c'$ either (i) $F(c,c) > F(c',c)$
or (ii) $F(c,c) = F(c',c)$ and $F(c,c') \ge F(c',c')$
\citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary dynamics can be applied to
meta games. We will later turn towards the replicator dynamic
\citep{TaylorJonker1978:Evolutionary-St} and the replicator mutator dynamic
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.\todo{footnote with caveat about agent-level
  processes}

@@

We submit that meta games are a useful tool for structuring and sharpening conceptual inquiry
in a domain that is notorious for its fuzziness and elusiveness. To show how this approach can
be applied to the benefit of conceptual debate, we consider the evolution of preference
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. Concretely, we here pit two different ways of
representing preferences against each other: (i) veridical representations that correctly
reflect objective fitness, and (ii) non-veridical subjective preferences in terms of regret
\citep[e.g.,][]{Savage1951:The-theory-of-s,}. We show that subjective, non-veridical preference
representations based on a notion of regret can be adaptive and even outperform objective,
veridical representations of the actual, true fitness, when agents, at least on occasion, hold
\emph{imprecise probabilistic beliefs} \citep[e.g.,][]{gilsch89,levi74,gardsah82} and make
decisions based on a security choice rule (here: the maxmin rule applied to expected
utility). This case study strips the generality of a meta game approach down to a perspicuous
comparison between only four choice mechanisms, summarized in Table~\ref{tab:CMs}: the
cross-product of two different preference representations (veridical vs.~regret) and two
different belief representations (precise vs.~imprecise).

\begin{table}[t]
  \centering
  \begin{tabular}{cccc}
    \multicolumn{2}{c}{subjective representation} & action selection & corresponding choice
    rule \\  \cmidrule(r){1-2}
    preference type & belief type  \\ \midrule
    objective fitness & precise & maxmin EU &  Laplace \\
    objective fitness & precise & maxmin EU & maxmin \\ 
    regret & imprecise & maxmin EU & Laplace \\
    regret & imprecise & maxmin EU & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent)
    classic choice rules. All notions are explained in the main text.}
  \label{tab:CMs}
\end{table}

This exclusive selection of choice mechanisms is motivated by a number of
considerations. Firstly, the set of conceivable choice mechanisms is vast. Most of that
conceivable vastness is utter nonsense. Some selection of plausible candidates is required in
any case. Secondly, a severe restriction to only a small number of suitable candidates will
help demonstrate more distinctly how a meta game approach may help structure philosophical
inquiry. This is our main goal. Thirdly, our case study adds to the recent literature on the
evolution of preferences by clearly demonstrating the possibly non-trivial interaction between
preference and belief representations, all else equal. This is our secondary goal. Finally, the
four choice mechanism that we look at here, although phrased in terms of preference and belief
representations, map onto well-known classic proposals for rational choice (see
Table~\ref{tab:CMs}). In other words, there is also an argument by historical relevance for our
selection. To see this, we first recapitulate the relevant classics ---to have a handy label,
we call them \emph{classic rules}--- and then show how they relate to our notion of choice
mechanism.

\subsection{Classic rules for rational choice} 
\label{sec:choice-rules}

The classic literature on rational choice contains a number of proposals for which acts a
player should choose if she is rational. Maximization of expected utilities, as
given in Definition~\ref{def:rationality}, is only one of them. Here, we consider three others:
(i) maxmin, (ii) Laplace's rule and (iii) regret minimization. All of these apply to games and
solitary decision situations. We focus on games, but definitions carry over straightforwardly.

A strategic game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a finite set of
players, $A_i$ a finite set of acts and $\pi_i$ a payoff function for each player $i \in N$.
Part of this paper will focus on symmetric games with two players. A two-player game is
symmetric if players have the same action set, $A_1 = A_2$, and payoffs are symmetric:
$\pi_1(a_k, a_j) = \pi_2(a_j, a_k)$ for all $a_j,a_k \in A_1$. An example is given in
Figure~\ref{coordgame1}. The numbers give the payoff for the player choosing a row act against
an opponent choosing a column act. Since the game is symmetric, it suffices to give just these
numbers.

Rational choice from maximization of expected utilities, as given in
Definition~\ref{def:rationality}, applies to games as follows. To say which acts are rational,
it requires specifying a belief of the DM. Let $A_{-i} = \bigtimes A_j$ for all $j \neq i$ be
the set of all \emph{profiles} of acts for all players except $i$. If player $i$ has a belief
$\mu_i \in \Delta(A_{-i})$ about the behavior of his opponents, he chooses rationally only if
his choice $a^*_i$ maximizes expected utility
$\EU(a_i) := \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. For
the example in Figure~\ref{coordgame1}, a rational player would choose act $I$ exactly when he
believes that his opponent will choose act $I$ with a probability of at least
$\nicefrac{2}{3}$.

Perhaps the most well-known classic rule that does not require explicitly stating the DM's
beliefs is the \emph{maxmin rule}. It tells our agent to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios:
$\act^* \in \argmax_{\myact{i} \in \Acts_i} \min_{\myact{-i} \in \Acts_{-1}} \pi_i(\myact{i},
\myact{-1})$.
For the example in Figure~\ref{coordgame1}, the maximim rule looks at the minima in each row,
asking: what is the worst possible outcome for each row act? It is $0$ in either
case. Consequently, the maxmin rule would not favor either act: players choosing by the maxmin
rule would be indifferent between acts $I$ and $II$.

\emph{Laplace's rule} selects $\act^*$ if
$\act^* \in \argmax_{\myact{-i} \in \Acts_{-i}} \sum_{\myact{-i}} \pi(\myact{i}, \myact{-i})$.
Laplace's rule is equivalent to expected utility maximization under a \emph{flat belief} that
assigns equal probability to all of the opponents' choices. In the coordination game of
Figure~\ref{coordgame1}, Laplace's rule selects act $II$.

The final classic rule that is relevant for our purposes here is \emph{regret
  minimization}. The notion of regret in decision theory dates back at least to the work by
\citet{Savage1951:The-theory-of-s}, and has later been developed by \citet{bell82}, \citet{fish82} and \citet{loosug82} independently. Lately,
\citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
solutions to game theoretical puzzles (like the Traveller's dilemma) in a way that is closer to
everyday intuition and empirical data. In this paper the notion of regret is defined as in
\citet{HalpernPass2012:Iterated-Regret}:

>>>>

\begin{abstract}
  Standard applications of evolutionary game theory look at a single, fixed game and focus on
  the evolution of behavior for that game alone. Instead, this paper uses tools from
  evolutionary game theory to study the evolutionary competition between \emph{choice
    mechanisms} in a rich and variable environment. A choice mechanism is a way of representing
  a decision situation, paired with a method for choosing an act based on this 
  subjective representation. We demonstrate the usefulness of this approach by a case study
  that compares a number of classic proposals of what it means to make a rational choice. Our
  main result is that regret-based preference representations can be evolutionarily selected
  for, despite being non-veridical.
\end{abstract}

@@

%Evolutionary game theory has become an established philosophical tool for probing into
%conceptual issues whose complexity requires mathematical modelling at a high level of
%abstraction.  Abstraction entails simplifying assumptions.

If agents deal with a rich and variable environment, they have to face many different choice
situations.
Standard evolutionary game models frequently
simplify reality in two ways that are relevant for present purposes. Firstly, the environment
is represented as a fixed \emph{stage game}; secondly, the focus of
evolutionary selection is behavior for that stage game alone. In contrast, some argue for
studying the evolutionary competition of general \emph{choice mechanisms} in a rich and
variable environment
\citep[e.g.,][]{FawcettHamblin2013:Exposing-the-be,McNamara2013:Towards-a-Riche}. To show that
there is no tension here, this paper introduces a method for reconciling established notions of
evolutionary game theory with evolutionary selection of choice mechanisms in variable
environments \citep[see related work
by][]{ZollmanSmead2010:Plasticity-and-,SmeadZollman2013:The-Stability-o}.\todo{More references
  Harley 1981!!!, O Conor \dots}

In full generality, a choice mechanism associates decision situations with action choices. A crucial
part of a choice mechanism is the \emph{subjective representation} of the decision situation,
in particular the manner of forming preferences and beliefs about a possibly uncertain world (see
Figure~\ref{fig:ChoiceMechanism} below). This paper asks: which preference and belief
representations are ecologically valuable and lead to high fitness?
The evolution of
preferences has been subject of recent interest in theoretical economics
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. A contribution of this paper is
to provide a case that questions of preference evolution should take
variability in uncertainty representation into account as well. As a point in case, we demonstrate
that if agents have imprecise probabilistic beliefs \citep[e.g.][]{gardsah82,levi74,walley96},
preference representations in terms of regret can be selected for, even though they may deviate
from the true, objective fitness that evolutionary selection operates on.

@@

Section \ref{sec:rati--subj} reviews different views of rational choice and relates the
approach we take here to the existing literature. Section \ref{sec:basic-notions} introduces the
relevant decision- and game-theoretic notions, and
%spells out four choice mechanisms on whose
%evolutionary competition we zoom in 
Section~\ref{sec:basic-model-1} presents the
main results of the paper. Section \ref{sec:extensions} extends the evolutionary analysis by
introducing more choice mechanisms and alternative models of the environment.

\section{Rationality, representations \& meta-games}
\label{sec:rati--subj}

The standard textbook definition of \textit{rationality} in economics and decision theory
traces back to the seminal work by \citet{deFinetti37}, \citet{Neumannvon-NeumannMorgenstern1944:Theory-of-Games}
and \citet{Savage1954:The-Foundations}. It says that a choice is rational only if it maximizes
(subjective) expected utility.

\begin{definition}[Rational choice from maximization of expected utility]
  \label{def:rationality}
  Let $S$ be a (finite) set of states and $A$ a (finite) set of actions. Given a
  probability measure $\mu$ over $S$ and a payoff function
  $u \mycolon S \times A \rightarrow \mathbb{R} $, an action $a^* \in A$ is rational
  if it maximizes the \emph{(subjective) expected utility}
  $\EU(a, \mu) := \sum_{s \in S} \mu(s) \ u(s, a)$.
\end{definition}

\noindent Expected utility is subjective in the sense that it is a function of subjective
beliefs $\mu$ and subjective preferences $u$ of the decision maker (DM). To wit, a choice can be
rational, i.e., the best choice from the DM's point of view, even if based on peculiar beliefs
and/or aberrant preferences. % (Indeed, while
% \citeauthor{Neumannvon-NeumannMorgenstern1944:Theory-of-Games} assumed that probabilities were
% objectively given, it was \citet{Savage1954:The-Foundations} who famously extended von Neumann
% and Morgenstern's axiomatization to include a subjective notion of probabilistic beliefs on top
% of the subjective notion of preference that von Neumann and Morgenstern started out with.)

If beliefs and preferences are subjective, there is room for \emph{rationalization} or
\emph{redescriptionism} of observable behavior. For example,
\citet{KahnemannTversky1979:Prospect-Theory} famously demonstrated that what appear to be
violations of rationality norms in human decision making can be explained on the assumption
that subjects' beliefs and preferences deviate systematically from the objectively given
parameters in the presented choice task. Similarly for social decision making, including social
preferences such as fairness, altruism or similar, allows us to describe as rational
empirically observed behavior, such as in experimental Prisoner's Dilemmas or public goods
games, that might otherwise appear irrational \citep[e.g.,][]{fehrschmidt99,charrab02}.

The main objection to redescriptionism is that, without additional constraints, the notion of
rationality is likely to collapse, as it seems possible to deem rational almost everything that
is observed, given the freedom to adjust beliefs and preferences at will. \emph{Normativism}
therefore emphasizes that there are many ways in which ascriptions of beliefs and preferences
should be constrained by normative considerations of rationality as well. As for beliefs, where
possible, subjective beliefs should reflect objective chance. 
%Where this is impossible,
%temporal consistency of choices constrains rational subjective beliefs, else exploitation by
%Dutch books is possible. 
As for preferences, in certain situations it may make sense to
postulate on normative grounds that subjective preferences should be oriented towards tracking
objective fitness. For instance, profit maximization seems a necessary requirement for
evolution in a competitive market because only firms behaving according to profit maximization will survive in the long run \citep[e.g.,][]{alch50,Fried53}.

An alternative view on rationality of choice is \emph{adaptationism}
\citep[e.g.,][]{Anderson1991:Is-human-cognit,ChaterOaksford2000:The-Rational-An,HagenChater2012:Decision-Making}. \todo{more
  references} Adaptationism aims to explain rational behavior by appeal to evolutionary
considerations: DMs have acquired choice mechanisms that have proved to be adpative with respect to the variable environment where they have evolved. A choice mechanism can be a set of distinct heuristics (the DM's
\emph{adaptive toolbox}) that have little in common
\citep[e.g.,][]{TverskyKahnemann1981:The-Framing-of-,GigerenzerGoldstein1996:Reasoning-the-F,ScheibehenneRieskamp2013:Testing-the-Ada}. But
to closely relate to the literature on evolution of preferences and to the philosophical debate about the nature of rational choice in general, we here suggest to
think of a choice mechanism as a map from choice situations to action choices which includes an
explicit level of subjective representation of the situation, that we call \emph{subjective representation scheme} (see
Figure~\ref{fig:ChoiceMechanism}). Specifically, a subjective representation scheme is a general way of forming preferences and
beliefs about the choice situation, and we are most interested in the question which
subjective representations are better than others from an evolutionary point of view.

@@

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = environment, node distance = 2cm] (situation)
    {environment produces random \\ \textbf{decision situation} $D$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = situation] (representation) 
    {\textbf{subjective representation} of $D$ \\ \textcolor{gray}{[preference, beliefs, \dots]}};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = representation] (action)
    {\textbf{action choice} $a$};

    \node[rounded corners, draw, rectangle, minimum height = 0.75cm, text width=6cm, text
    centered, below of = action,node distance = 1.5cm] (payoff) {\textbf{payoff} for $a$ in $D$};

@@

\end{tikzpicture}
  \caption{Schematic representation of action choice. A choice mechanism maps decision
    situation $D$ to action choice $a$ by means of a representation of $D$ in terms of a subjective
    preference and a subjective belief.}
  \label{fig:ChoiceMechanism}
\end{figure}

The fitness of a subjective representation scheme is measured in terms of the fitness of a
choice mechanism that contains it, all else equal. The fitness of a choice mechanism $c$ is the
expected payoff of $c$ over arbitrary decision situations $D$:
\begin{align}
  \label{eq:FittnessChoiceMechSolitary}
  F(c) = \int P(D) \  \pi_D(c) \text{ d} D \,,
\end{align}
where $P(D)$ is the probability (density) of $D$ occurring and $\pi_D(c)$ is the evolutionary payoff of
the act selected by $c$ in $D$. The probability measure $P(D)$ encodes the statistical
properties of the environment.

The same logic also applies to game contexts and social decision making. A game $G$ is also a
choice situation, albeit a more complex one. A choice mechanism $c$ maps $G$ onto an action
choice in $G$. The fitness of $c$ is then:\footnote{We assume for simplicity of exposition that
  all games with positive occurrence probability are symmetric two-player games.}
\begin{align}
  \label{eq:FittnessChoiceMechGame}
  F(c) = \int \int P(G) \  P(c') \  \pi_G(c,c') \text{ d} c' \text{ d} G  \,,
\end{align}
where $P(G)$ is the probability (density) of game $G$, $P(c')$ is the probability of
encountering a co-player who uses choice mechanism $c'$ and $\pi_G(c,c')$ is the evolutionary payoff for
a $c$-player in $G$ against the $c'$-player, given by the actions that $c$
and $c'$ select in $G$. For the most part, we will focus on the more complex case of game situations. Results about solitary decision making follow from parallel arguments
(see Section~\ref{sec:extensions}).

\emph{Meta-games} are abstract models for the evolutionary competition between choice
mechanisms in interactive decision making contexts. The expected payoff for a $c$-player in a random
encounter with a $c'$-player is the average of payoffs obtained in a random symmetric game $G$ weighted
by $G$'s occurrence probability:
\begin{align}
  \label{eq:FittnessChoiceMechGamePairwise}
  F(c, c') = \int P(G) \  \pi_G(c,c') \text{ d} G \,.
\end{align}

\begin{definition}[Meta-game]
  \label{def:MetaGame}
  Let $CM$ be a set of choice mechanisms and let $P(G)$ give the occurrence probability of symmetric game
  $G$ in the environment. The meta-game that captures selection of choice mechanisms in this
  environment is the one-population evolutionary game with strategies $CM$ and payoff function
  $\pi \mycolon CM \times CM \rightarrow \mathbb{R}$ such that
  $\pi(c,c') = F(c,c')$.
  % = \int P(G) \  \pi_G(c,c') \text{ d} G$.
\end{definition}

Standard notions of evolutionary game theory can be applied to meta-games as well. A choice
mechanism $c$ is a strict Nash equilibrium if $F(c,c) > F(c',c)$ for all $c'$; it is
evolutionarily stable if for all $c'$ either (i) $F(c,c) > F(c',c)$ or (ii) $F(c,c) = F(c',c)$
and $F(c,c') > F(c',c')$; it is neutrally stable if for all $c'$ either (i) $F(c,c) > F(c',c)$
or (ii) $F(c,c) = F(c',c)$ and $F(c,c') \ge F(c',c')$
\citep{Maynard-Smith1982:Evolution-and-t}. Similarly, evolutionary dynamics can be applied to
meta-games. We will later turn towards replicator dynamics
\citep{TaylorJonker1978:Evolutionary-St} and replicator mutator dynamics
\citep[e.g.][]{Nowak2006:Evolutionary-Dy}.
%\todo{footnote with caveat about agent-level processes}

@@

We submit that meta-games are a useful tool for structuring and sharpening conceptual inquiry
in a domain that is notorious for its fuzziness and elusiveness. To show how this approach can
be applied to the benefit of conceptual debate, we consider the evolution of preferences
\citep[e.g.,][]{algweib13,DekElyYlan07,RobSam11}. 
%Concretely, we here pit two different ways of
%representing preferences against each other: (i) veridical representations that correctly
%reflect objective fitness, and (ii) non-veridical subjective preferences in terms of regret
%\citep[e.g.,][]{Savage1951:The-theory-of-s,}. 
We show that subjective, non-veridical preference
representations based on a notion of regret can be adaptive and even outperform objective representations of the true evolutionary fitness, when agents, at least on occasion, hold
\emph{imprecise probabilistic beliefs} \citep[e.g.,][]{gilsch89,levi74,gardsah82} and make
decisions based on a security choice rule (here: the maxmin rule).\footnote{Decision-theoretic background and justifications for the use of maxmin and imprecise probabilities is given in Appendix~\ref{sec:impr-prob-beli}.}

The case study under consideration strips the generality of a meta-game approach down to a perspicuous
comparison between four choice mechanisms, summarized in Table~\ref{tab:CMs}.
% the cross-product of two different preference representations (veridical vs.~regret) and two
%different belief representations (precise vs.~imprecise). 
\begin{table}[t]
  \centering
  \begin{tabular}{cccc}
  \multicolumn{3}{c}{choice mechanism} \\  \cmidrule(r){1-3}
    \multicolumn{2}{c}{subjective representation} & action selection & corresponding classic rule \\  \cmidrule(r){1-2}
    preference type & belief type  \\ \midrule
    objective fitness & precise & maxmin EU &  Laplace rule \\
    objective fitness & imprecise & maxmin EU & maxmin \\ 
    regret & precise & maxmin EU & Laplace rule \\
    regret & imprecise & maxmin EU & regret minimization \\ 
  \end{tabular}
  \caption{Overview relevant choice mechanisms and the corresponding (behaviorally equivalent)
    classic choice rules.}
  \label{tab:CMs}
\end{table}
\iffalse
This exclusive selection of choice mechanisms is motivated by a number of
considerations. Firstly, the set of conceivable choice mechanisms is vast. Most of that
conceivable vastness is utter nonsense, and some selection of plausible candidates is required in
any case. Secondly, a restriction to only a small number of suitable candidates will
help demonstrate more distinctly how a meta game approach may help structure philosophical
inquiry, which is our main goal. Thirdly, our case study adds to the recent literature on the
evolution of preferences by making evident the possibly non-trivial interaction between
preference and belief representations, all else equal. This is our secondary goal. 
Finally,
\fi 
The
four choice mechanisms,  
%that we look at here
although phrased in terms of preference and belief
representations, map onto well-known classic proposals for rational decision making (see
Table~\ref{tab:CMs}). In other words, there is also an argument by historical relevance for our
selection. To see this, we first recapitulate the relevant classics ---to have a handy label,
we call them \emph{classic rules}--- and then show how they relate to our notion of choice
mechanism.

\subsection{Classic rules for decision making}
\label{sec:choice-rules}

The classic literature on rational choice contains a number of proposals for which acts a
player should choose if she is rational. Maximization of expected utility, as
given in Definition~\ref{def:rationality}, is just one of them, possible only if DM holds a precise probabilistic belief.
Here, we consider three others:
(i) maxmin, (ii) Laplace rule, and (iii) regret minimization. 
%All of these apply to games and
%solitary decision situations.
We focus on games first, but definitions carry over straightforwardly to solitary decisions.

A strategic-form game is a tuple $ G=\tuple{ N, (A_i , \pi_i)_{i \in N} }$ with $N$ a finite set of
players, $A_i$ a finite set of acts and $\pi_i: \bigtimes_{j \in N} A_j \rightarrow \mathbb{R}$ a payoff function for each player $i \in N$.
Part of this paper will focus on symmetric games with two players. A two-player game is
symmetric if players have the same action set, $A_1 = A_2$, and payoffs are symmetric:
$\pi_1(a_k, a_j) = \pi_2(a_j, a_k)$ for all $a_j,a_k \in A_1$. An example is given in
Figure~\ref{coordgame1}. The numbers give the payoff for the player choosing a row act against
a co-player choosing a column act. Since the game is symmetric, it suffices to give just these
numbers.

Rational choice from maximization of expected utility, as given in
Definition~\ref{def:rationality}, applies to games as follows. To say which acts are rational,
it requires specifying a precise probabilistic belief of DM. Let $A_{-i} := \bigtimes_{j \neq i} A_j$ be
the set of all \emph{profiles} of acts for all players except $i$. If player $i$ has a belief
$\mu_i \in \Delta(A_{-i})$ about the behavior of his co-players, he chooses rationally only if
his choice $a^*_i$ maximizes expected utility
$\EU(a_i, \mu) := \sum_{\act_{-i} \in \Acts_{-i}} \mu_i(\act_{-i}) \myts \pi_i(a_i, a_{-i})$. For
the example in Figure~\ref{coordgame1}, an expected utility maximizer would choose act $I$ only if he
believes that his co-player is going to choose act $I$ with a probability of at least
$\nicefrac{2}{3}$.

Perhaps the most well-known classic rule that does not require precise probabilistic
beliefs is the \emph{maxmin rule}. It prescribes to choose act $\act^*$ only if it
maximizes payoffs for all worst case scenarios:
$\act^* \in \argmax_{a_i \in \Acts_i} \min_{a_{-i} \in \Acts_{-i}} \pi_i(a_{i},
a_{-i})$.
For the example in Figure~\ref{coordgame1}, the minimum of both acts is 0: players choosing by maxmin
rule would be indifferent between acts $I$ and $II$.

\emph{Laplace rule} selects $\act^*$ if
$\act^* \in \argmax_{a_{i} \in \Acts_{i}} \sum_{a_{-i}} \pi(a_{i}, a_{-i})$.
Laplace rule is equivalent to expected utility maximization under a \emph{flat belief} that
assigns equal probability to all of the co-players' choices. In the coordination game of
Figure~\ref{coordgame1}, Laplace rule selects act $II$.

The final classic rule that is relevant for our purposes is \emph{regret
  minimization}. The notion of regret in decision theory dates back at least to the work by
\citet{Savage1951:The-theory-of-s}, and has later been developed by \citet{bell82}, \citet{fish82} and \citet{loosug82} independently. Lately,
\citet{HalpernPass2012:Iterated-Regret} showed how the use of regret minimization can give
solutions to game theoretical-puzzles (like the Traveller's dilemma and the Centipede game) in a way that is closer to
everyday intuition and empirical data. In this paper the notion of regret is defined as in
\citet{HalpernPass2012:Iterated-Regret}:
